Domain-agnostic Outlier Ranking Algorithms (DORA)---A Configurable Pipeline for Facilitating Outlier Detection in Scientific Datasets

### 2022-01-11 publications/frontiers/dora_outlier_detection_frontiers.tex
Fill in todos. #79

% packages added by authors
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage{amsmath}
\usepackage[dvipsnames]{xcolor}
\newcommand{\todo}[1]{\textcolor{blue}{#1}}

@@

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Kerner {et~al.}} %use et al only if is more than 1 author
\def\Authors{Hannah R. Kerner\,$^{1,*}$, Umaa Rebbapragada\,$^{2}$, 
Kiri L. Wagstaff\,$^{2}$, Steven Lu\,$^{2}$, Bryce Dubayah\,$^{1}$,
Eric Huff\,$^{2}$, Jake Lee\,$^{2}$, Vinay Raman\,$^{3}$, and
Sakshum Kulshrestha\,$^{1}$}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{$^{1}$Department of Geographical Sciences, University of Maryland, College Park, MD, USA \\
$^{2}$Jet Propulsion Laboratory, California Institute of Technology, Pasadena, California, USA \\
$^{3}$Montgomery Blair High School, Silver Spring, MD, USA}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Corresponding Author}

@@

\begin{table}
\caption{Number of samples in the training and test sets for each dataset,
ordered by dataset size.}
\label{tab:datasets}
\centering
\begin{tabular}{l|c|cc}
  \hline
  Dataset & Training & \multicolumn{2}{c}{Test}\\
   & Unlabeled &  Outliers &  Inliers \\

  \hline
  Planetary & 992 &  9 & 49 \\
  Earth & 6,757 & 37 & 76  \\
  F-MNIST & 60,000 & 1,000 & 1,000  \\
  Astrophysics & 100,000   &  25,339 &  74,661 \\
  \hline
\end{tabular}
\end{table}

@@

\begin{table}
\caption{Precision at $N=n_{outlier}$ for four datasets; the best result for each data set is in bold.}
\label{tab:faw_results}
\centering
\begin{tabular}{l|cccc}
  \hline
  Algorithm & Astro. & Earth & Planetary & F-MNIST \\
  \hline
  PCA               & \textbf{0.42} & 0.41 & 0.44 & \textbf{0.84} \\
  DEMUD             & ---- & 0.41 & 0.44 & \textbf{0.84} \\
  \hline
  RX                & 0.40 & 0.43 & ---- & 0.82 \\
  LRX               & ----& ---- & 0.33 & 0.56 \\
  \hline
  Iso. Forest  & 0.34 & 0.46 & \textbf{0.56} & 0.74 \\
  \hline
  PAE               & 0.35 & 0.41 & 0.44 & 0.83 \\
  Neg. Sampling & 0.32 & \textbf{0.49} & 0.33 & 0.43 \\
  \hline
  Random            & 0.25 & 0.32 & 0.14 & 0.50 \\
  \hline
\end{tabular}
\end{table}

@@

% This section was specific to IAAI, can change the name, remove, or keep
% as is if we want.
\subsection{Path to deployment} 
Our goal is for DORA to enable increased application and benefit of
outlier detection methods in real-world scientific use cases. We have 
designed DORA to make it as easy as possible for scientists to apply
algorithms and to compare and interpret their results; users need only
to specify the specifics of their data (e.g., path, data type) in a 
configuration file to start running experiments and seeing results. We will
make DORA publicly available as a pip-installable package to make it
easy to integrate into existing scientific workflows. 
The source code and example datasets used in this study will also be 
open-sourced, enabling DORA to be improved and expanded by the 
machine learning and domain science communities. 
In addition, DORA will be infused into the scientific workflows for the
three use cases we demonstrated results for in this study.
\todo{Add link to github since open now.}
%the use cases from this will also lead to infusion in real scenarios

@@

\section*{Author Contributions}

The Author Contributions section is mandatory for all articles, including articles by sole authors. If an appropriate statement is not provided on submission, a standard one will be inserted during the production process. The Author Contributions statement must describe the contributions of individual authors referred to by their initials and, in doing so, all authors agree to be accountable for the content of the work. Please see  \href{http://home.frontiersin.org/about/author-guidelines#AuthorandContributors}{here} for full authorship criteria. \todo{Fill this in.}

\section*{Funding}
We thank the NASA Center Innovation Fund and the NASA Science Mission 
Directorate (under the NASA Cross-Divisional AI Use Case Demonstration grants)
 for supporting this work. \todo{Add grant numbers if possible.}

\section*{Acknowledgments}
We thank the UN Food and Agriculture Organization (FAO) and Dr. Ritvik Sahajpal
(University of Maryland/NASA Harvest) for providing the maize ground
 observation dataset.

 We thank Raymond Francis for analyzing the Navcam rover images to
identify targets considered to be ``novel'' for evaluating outlier
detection in the planetary rover scenario.

Part of this research was carried out at the Jet Propulsion
Laboratory, California Institute of Technology, under a contract with
the National Aeronautics and Space Administration.

@@

\section*{Data Availability Statement}
The datasets [GENERATED/ANALYZED] for this study can be found in the [NAME OF REPOSITORY] [LINK]. 
\todo{Can we make our datasets available here? We can at least add the github link to the code.}
% Please see the availability of data guidelines for more information, at https://www.frontiersin.org/about/author-guidelines#AvailabilityofData

>>>>

% packages added by authors
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage[dvipsnames]{xcolor}
\newcommand{\todo}[1]{\textcolor{blue}{#1}}

@@

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Kerner {et~al.}} %use et al only if is more than 1 author
\def\Authors{Hannah R. Kerner\,$^{1,*}$, Umaa Rebbapragada\,$^{2}$, 
Kiri L. Wagstaff\,$^{2}$, Steven Lu\,$^{2}$, Bryce Dubayah\,$^{1}$,
Eric Huff\,$^{2}$, Jake Lee\,$^{2}$, Vinay Raman\,$^{3}$, and
Sakshum Kulshrestha\,$^{1}$}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{$^{1}$University of Maryland, College Park, MD, USA \\
$^{2}$Jet Propulsion Laboratory, California Institute of Technology, Pasadena, California, USA \\
$^{3}$Montgomery Blair High School, Silver Spring, MD, USA}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Corresponding Author}

@@

\begin{table}
\caption{Number of samples in the training and test sets for each dataset,
ordered by dataset size.}
\label{tab:datasets}
\centering
\begin{tabular}{l|c|cc}
  \toprule
  Dataset & Training & \multicolumn{2}{c}{Test}\\
   & Unlabeled &  Outliers &  Inliers \\

  \midrule
  Planetary & 992 &  9 & 49 \\
  Earth & 6,757 & 37 & 76  \\
  F-MNIST & 60,000 & 1,000 & 1,000  \\
  Astrophysics & 100,000   &  25,339 &  74,661 \\
  \bottomrule
\end{tabular}
\end{table}

@@

\begin{table}
\caption{Precision at $N=n_{outlier}$ for four datasets; the best result for each data set is in bold.}
\label{tab:faw_results}
\centering
\begin{tabular}{l|cccc}
  \toprule
  Algorithm & Astro. & Earth & Planetary & F-MNIST \\
  \midrule
  PCA               & \textbf{0.42} & 0.41 & 0.44 & \textbf{0.84} \\
  DEMUD             & ---- & 0.41 & 0.44 & \textbf{0.84} \\
  \midrule
  RX                & 0.40 & 0.43 & ---- & 0.82 \\
  LRX               & ----& ---- & 0.33 & 0.56 \\
  \midrule
  Iso. Forest  & 0.34 & 0.46 & \textbf{0.56} & 0.74 \\
  \midrule
  PAE               & 0.35 & 0.41 & 0.44 & 0.83 \\
  Neg. Sampling & 0.32 & \textbf{0.49} & 0.33 & 0.43 \\
  \midrule
  Random            & 0.25 & 0.32 & 0.14 & 0.50 \\
  \bottomrule
\end{tabular}
\end{table}

@@

% This section was specific to IAAI, can change the name, remove, or keep
% as is if we want.
\subsection{Path to deployment} 
Our goal is for DORA to enable increased application and benefit of
outlier detection methods in real-world scientific use cases. We have 
designed DORA to make it as easy as possible for scientists to apply
algorithms and to compare and interpret their results; users need only
to specify the specifics of their data (e.g., path, data type) in a 
configuration file to start running experiments and seeing results. We will
make DORA publicly available as a pip-installable package to make it
easy to integrate into existing scientific workflows. 
The source code and example datasets used in this study will also be 
open-sourced, enabling DORA to be improved and expanded by the 
machine learning and domain science communities. 
In addition, DORA will be infused into the scientific workflows for the
three use cases we demonstrated results for in this study.
The DORA code is openly accessible via Github at
\url{https://github.com/nasaharvest/dora}.
%the use cases from this will also lead to infusion in real scenarios

@@

\section*{Author Contributions}
HRK, UR, and KLW developed the conception of the project. HK oversaw the overall implementation and led the analysis of the Earth science dataset. UR led the analysis of the astrophysics dataset. KLW led the analysis of the planetary science dataset. SL led the development of the DORA software architecture. BD led the analysis of the FMNIST/MNIST dataset. EH supported the analysis of the astrophysics dataset. JL developed the desktop application for results visualization. VR supported the analysis of the Earth science dataset. SK supported the analysis of the planetary science dataset. All authors contributed to the overall system implementation and manuscript writing/editing.

\section*{Funding}
We thank the NASA Center Innovation Fund (\todo{Award number?})
and the NASA Science Mission Directorate 
(under the NASA Cross-Divisional AI Use Case Demonstration grants;
Award \#80NSSC21K0720) for supporting this work.

\section*{Acknowledgments}
We thank the UN Food and Agriculture Organization (FAO) and Dr. Ritvik Sahajpal
(University of Maryland/NASA Harvest) for providing the maize ground
 observation dataset. 
%
We thank Raymond Francis for analyzing the Navcam 
rover images to identify targets considered to be ``novel'' for evaluating
outlier detection in the planetary rover scenario.
%
Part of this research was carried out at the Jet Propulsion
Laboratory, California Institute of Technology, under a contract with
the National Aeronautics and Space Administration.

@@

\section*{Data Availability Statement}
The datasets [GENERATED/ANALYZED] for this study can be found in the [NAME OF REPOSITORY] [LINK]. 
% Please see the availability of data guidelines for more information, at https://www.frontiersin.org/about/author-guidelines#AvailabilityofData
